\documentclass[paper=8.27in:11.69in, 14pt, DIV=calc]{scrartcl}
\usepackage{geometry}
\usepackage{graphics,graphicx}
\usepackage{pdfpages}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{minibox}
\newcounter{numbers}
\newcommand\printnumbers{\refstepcounter{numbers}\thenumbers}
\newcounter{answers}
\newcommand\printanswers{\refstepcounter{answers}\theanswers}

\begin{document}

\textbf{\begin{center}
\begin{Large}
CSCI 6370 IR and Web Search\\
ASSIGNMENT 2\\
Due is 06/15/2020 23:59\\
Ulvi Bajarani\\
Student ID 20539914\\
E-mail: ulvi.bajarani01@utrgv.edu\\
\end{Large}
\end{center}}

\newpage
\noindent \begin{center}
\textbf{Questions and Answers:}
\end{center}

\textbf{Problem \printnumbers . This assignment is designed for you to get familiar with various evaluation mechanisms in IR. Given the set of data in the table,}

\begin{enumerate}
\item \textbf{Compute the recall/precision rate at each relevant document point.}
\item \textbf{Compute F-Measure at each  relevant document point.}
\item \textbf{Compute the corresponding E-measure with $\beta = 0.8$ at each relevant document point.}
\item \textbf{Compute the R-Precision.}
\item \textbf{Computer the average precision, which is the average of the precision values at the points at which each relevant document is retrieved.}
\end{enumerate}

\textbf{The total relevant documents in a document set is 5.  The top 10 retrieved documents are
listed with the ones that are relevant marked.}

\begin{tabular}{|c|c|c|}
\hline
\textbf{Order} & \textbf{Doc \#} & \textbf{Relevant} \\ \hline
1              & 586             & X                 \\ \hline
2              & 357             &                   \\ \hline
3              & 358             & X                 \\ \hline
4              & 108             & X                 \\ \hline
5              & 345             &                   \\ \hline
6              & 114             &                   \\ \hline
7              & 555             & X                 \\ \hline
8              & 888             &                   \\ \hline
9              & 860             &                   \\ \hline
10             & 167             & X                 \\ \hline
\end{tabular}
\\
\textbf{\\Answer \printanswers .\\}
\begin{enumerate}
\item The answer is provided in the tables.
\item The answer is provided in the tables.
\item The answer is provided in the tables.
\item R-Precision = $P(5)$ = 0.6.
\item MAP = 0.816706349206349
\end{enumerate}

\begin{tabular}{|c|c|c|c|c|c|}
\hline
\textbf{Order} & \textbf{Doc \#} & \textbf{Relevant} & Number of found documents & \textbf{Recall} & \textbf{Precision} \\ \hline
1  & 586 & X & 1 & 0.2 & 1                 \\ \hline
2  & 357 &   & 1 &     &                   \\ \hline
3  & 358 & X & 2 & 0.4 & 0.666666666666667 \\ \hline
4  & 108 & X & 3 & 0.6 & 0.75              \\ \hline
5  & 345 &   & 3 &     &                   \\ \hline
6  & 114 &   & 3 &     &                   \\ \hline
7  & 555 & X & 4 & 0.8 & 0.571428571428571 \\ \hline
8  & 888 &   & 4 &     &                   \\ \hline
9  & 860 &   & 4 &     &                   \\ \hline
10 & 167 & X & 5 & 1   & 0.5               \\ \hline
\end{tabular}
\\
\\
\\
\\
\begin{tabular}{|c|c|c|c|c|c|}
\hline
\textbf{Order} & \textbf{Doc \#} & \textbf{Relevant} & Number of found documents & \textbf{F-Measure} & \textbf{E-measure (b=0.8)} \\ \hline
1  & 586 & X & 1 & 0.333333333333333 & 0.609523809523809 \\ \hline
2  & 357 &   & 1 &                   &                   \\ \hline
3  & 358 & X & 2 & 0.5               & 0.470967741935484 \\ \hline
4  & 108 & X & 3 & 0.666666666666667 & 0.316666666666667 \\ \hline
5  & 345 &   & 3 &                   &                   \\ \hline
6  & 114 &   & 3 &                   &                   \\ \hline
7  & 555 & X & 4 & 0.666666666666667 & 0.356862745098039 \\ \hline
8  & 888 &   & 4 &                   &                   \\ \hline
9  & 860 &   & 4 &                   &                   \\ \hline
10 & 167 & X & 5 & 0.666666666666667 & 0.378787878787879 \\ \hline
\end{tabular}

\textbf{Problem \printnumbers . Suppose we use the following the method to reformulate the query vector in response to relevance feedback:\\}

$q_{m} = q_{0} +  \sum_{d \ \text{is relevant}} d - \sum_{d' \ \text{is irrelevant}} d'$\\

\textbf{Consider the initial query vector is $q_{0} \ = \ \{1, 2, 0, 4, 0, 1\}$. The relevant feedback gives two relevant vectors $d_{1}$ and $d_{2}$ and one irrelevant vector $d_{3}$ as follows:\\}

$d_{1} \ = \ \{0, 1, 1, 2, 0, 2\}$
$d_{2} \ = \ \{5, 0, 2, 0, 2, 0\}$
$d_{3} \ = \ \{4, 2, 1, 2, 1, 3\}$

\textbf{\\Calculate the reformulated query vector $q_{1}$\\}.
\\
\textbf{Answer \printanswers .}
\\
$q_{m} \ = \ \{1+(0+5)-4 ; 2+(1+0)-2 ; 0+(1+2)-1 , 4+(2+0)-2, 0+(0+2)-1, 1+(2+0)-3\}$
$q_{m} \ = \ \{2;1;2;4;1;0\}$\\
\\
\textbf{Problem \printnumbers . Phrasal query is to retrieve documents with a specific phrase (ordered list of contiguous words). For example, phrasal query for  Q=``computer learning theory'' needs to retrieve all documents containing the phrase ``computer learning theory''. We usually ignore cases of the letters in the query. We may also allow intervening stop words and/or stemming. Describe an algorithm to do phrasal query with help of an inverted index.\\}
\\
\textbf{\\Answer \printanswers .\\}
Find set of documents $D$ in which all keywords ($k_{1}$...$k_{m}$) in phrase occur (using AND query processing).\\
Intitialize empty set, $R$, of retrieved documents.\\
For each document, $d$, in $D$:\\
\indent Get array, $P_{i}$ ,of positions of occurrences for each $k_{i}$ in $d$\\
\indent Find shortest array $P_{s}$ of the $P_{i}$’s\\
\indent For each position $p$ of keyword $k_{s}$ in $P_{s}$\\
\indent \indent For each keyword $k_{i}$ except $k_{s}$\\
\indent \indent Use binary search to find a position (p – s + i) in the array $P_{i}$\\
\indent If correct position for every keyword found, add $d$ to $R$\\
Return $R$\\

\end{document}